The Ex-Ante View of Recommender System Design

### 2019-04-20 paper.tex

Recommender Systems (RS) have been widely deployed in online platforms such as Netflix, Amazon, and YouTube. On these platforms the combination of the fact that most items are experience goods (i.e. valuation is only learned after consumption) and that the number of choices available to users is incredibly large, results in users not knowing their true preference ordering over all of the items. For instance, on Netflix there are thousands of movies, on Amazon millions of products, and YouTube billions of videos. As a result, RS have been useful in helping users make decisions on these platforms because they provide users with information to help them make their choices under uncertainty. 
\par
However, while acknowledging that this is why RS are useful, most of the literature on RS ignores this user uncertainty and takes what we call an \textit{ex-post} viewpoint to preferences, focusing on the reported utility of users after they consume the item or inferring them via behavioral data \cite{adomavicius2005toward, zhao2018interpreting}. The RS problem traditionally is cast as solving the following prediction problem:
\begin{align*}
\arg\max\limits_{x \in A} u(x)
\end{align*}
\noindent where $A$ is the set of possible items and $u(x)$ represents the utility a user gets to consume good $x$. Recently there has been a move away from metrics focused on maximizing the accuracy of this prediction problem and towards optimizing for metrics such as serendipity or discoverability \cite{mcnee2006being, vargas2011rank} due to  the realization that accurate recommendations may not be the most useful recommendations for users. Furthermore, deployed RS have come under scrutiny for leading to ``filter bubbles" \cite{pariser2011filter} where users consume a less diverse set of content over time as well as increased homogenization amongst users \cite{chaney2018algorithmic, hosanagar2013will}.
\par
In this paper we want to put forward the argument that one approach to overcoming these problems is to make understanding how users choose which items to consume a first-order consideration when designing RS. In particular, we differentiate between the previous \textit{ex-post} interpretation of the RS problem and what we define as the \textit{ex-ante} viewpoint. Unlike the ex-post viewpoint taken in the traditional prediction problem, the ex-ante viewpoint simply states that users face uncertainty and so maximize their expected utility as is defined in economic decision theory under uncertainty \citep[see e.g.][]{mas1995microeconomic}:
\begin{align*}
\arg\max\limits_{x \in A} \mathbb{E}[u(x)]
\end{align*}
Fundamentally, the move away from accuracy towards alternative metrics as well as tensions that have arisen from increased user homogenization and filter bubbles come from the \textit{choices} that users make. This shift from thinking about true preferences after all uncertainty is resolved (the ex-post view) to user decisions under uncertainty (the ex-ante view) can serve as a theoretical tool to better understand what recommendations are useful for users as well as understanding what of the documented perverse effects of RS come from recommendation per-se and what effects come from inherent uncertainty in user decision-making.
\section{The Value of the Ex-Ante View}

>>>>

Recommender Systems (RS) have been widely deployed in online platforms such as Netflix, Amazon, and YouTube. On these platforms the combination of most goods being experience goods and the number of choices available to users being incredibly large, results in users not knowing their true preference ordering over all of the items. For instance, on Netflix there are thousands of movies, Amazon millions of products, and YouTube billions of videos. As a result, RS have been useful in helping users make decisions on these platforms because they provide users with information to help them make their choices under uncertainty. 
\par
However, while acknowledging that this is why RS are useful, most of the literature on RS ignores this user uncertainty and takes what we call an \textit{ex-post} viewpoint to preferences, focusing on the reported utility of users after they consume the item or inferring them via behavioral data \cite{adomavicius2005toward, zhao2018interpreting}. The RS problem traditionally is cast as solving the following prediction problem:
\begin{align*}
\arg\max\limits_{x \in A} u(x)
\end{align*}
\noindent where $A$ is the set of possible items and $u(x)$ represents the utility a user gets to consume good $x$. Recently there has been a move away from metrics focused on maximizing the accuracy of this prediction problem and towards optimizing for metrics such as serendipity or discoverability \cite{mcnee2006being, vargas2011rank} due to  the realization that accurate recommendations may not be the most useful recommendations for users. Furthermore, deployed RS have come under scrutiny for leading to ``filter bubbles" \cite{pariser2011filter} where users consume a less diverse set of content over time as well as increased homogenization amongst users \cite{chaney2018algorithmic, hosanagar2013will}.
\par
In this paper we want to put forward the argument that one approach to overcoming these problems is to make understanding how users choose which items to consume a first-order consideration when designing RS. In particular, we differentiate between the previous \textit{ex-post} interpretation of the RS problem and what we define as the \textit{ex-ante} viewpoint. Unlike the ex-post viewpoint taken in the traditional prediction problem, the ex-ante viewpoint simply states that users face uncertainty and so maximize their expected utility as is defined in economic decision theory under uncertainty \citep[see e.g.][]{mas1995microeconomic}:
\begin{align*}
\arg\max\limits_{x \in A} \mathbb{E}[u(x)]
\end{align*}
Fundamentally, the move away from accuracy towards alternative metrics as well as tensions that have arisen from increased user homogenization and filter bubbles come from the \textit{choices} that users make. This shift from thinking about true preferences after all uncertainty is resolved (the ex-post view) to user decisions under uncertainty (the ex-ante view) can serve as a theoretical tool to better understand what recommendations are useful for users as well as understanding what of the documented perverse effects of RS come from recommendation per-se and what effects come from inherent uncertainty in user decision-making.
\section{The Value of the Ex-Ante View}
